---
title: "Covariance Model for Teleconnection"
output:
  pdf_document: default
  html_notebook: default
---

# 1. Covariance Model

## 1.1 nearest neighbour and teleconnection

```{r,include=FALSE}
library(MASS)
library(Matrix)
p=30
Ip=diag(rep(1,p))
V=matrix(0,p,3)
r=0.5
s=c(r,1,r)
p1=5
V[(p1-1):(p1+1),1]=s
p1=27
V[(p1-1):(p1+1),1]=s
p1=7
V[(p1-1):(p1+1),3]=s
p1=9
V[(p1-1):(p1+1),2]=s


S=V%*%t(V)
```

Consider the matrix $V$:

```{r, fig.width=4, fig.height=4, echo=FALSE}
print(V)
image(as(V, "dgCMatrix"),main="V")
```

Let the covariance matrix be $$S=VV^\top+I$$
```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(V%*%t(V)+Ip, "dgCMatrix"),main="S=VV'+I")
```


Meanwhile the inverse of $S$ is


```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(solve(S+Ip), "dgCMatrix"),main="S inverse")
```


## 1.2 link with factor model


$$
X=\begin{bmatrix}
    X_{1}   \\
    X_{2}    \\
    \vdots \\
    X_{p}
\end{bmatrix},
F=\begin{bmatrix}
    f_{1}   \\
    f_{2}    \\
    \vdots \\
    f_{m}
\end{bmatrix}
$$
$$X=\mu+AF+\epsilon$$
$$\Sigma=cov(X)=AA^\top+I$$

$A$ is not unique. Change the model a little bit, let $$V=[v_1,\cdots,v_r]$$, and $$\Lambda=diag(\lambda_1,\cdots,\lambda_r)$$. Then we have

$$\Sigma=V\Lambda V^\top+I=\sum_{i=1}^{r}\lambda_iv_iv_i^\top+I$$.

Note that each column of $V$: coefficients for certain factor for different features $X$

We have:

- Suppose: features $X_i$ is indexed by $i$, associated with the distance
- neighbourhood effect: row vector similarity, i.e. the more similar, the more correlated, 
- teleconnection effect: remote $i$ and $j$ are driven by the same factor, i.e. the column of $V$ (the common factor) has two separate (remote) peaks

Possible model assumption:

- If column of $V$, i.e. $v_i$ is **single peaked**, then it can be used to model neighbourhood effect
- If $v_i$ is **double peaked**, then it can be used to model teleconnection
- We can use kernel function to model the peak.
- We can assume the support of the kernel (or the peak) is sparse, since not all the features can be related to the same factor.

For example,

- Let $K(\cdot)$ be a kernel function
- $v_i$ has peak center $p_i$
- $v_i=K(\frac{i-p_i}{h})$, where $h$ is the bandwidth
- We can assume one teleconnection, for example, $v_1=K(\frac{i-p^{(1)}_i}{h})+K(\frac{i-p^{(2)}_i}{h})$
- $\Sigma=V\Lambda V^\top+I=\sum_{i=1}^{r}\lambda_iv_iv_i^\top+I$

In this way,$\Sigma$ is under a spiked covariance model with different sparsity structure of the spikes ($v_i$): **the covariance is the sum of many single peaked and one double peaked vectors**. 

# 2. Different Sparsity

## 2.1 group sparsity

```{r,include=FALSE}
p=30
Ip=diag(rep(1,p))
V=matrix(0,p,3)
p1=5
V[p1,]=rnorm(3)
p1=27
V[p1,]=rnorm(3)
p1=7
V[p1,]=rnorm(3)
p1=9
V[p1,]=rnorm(3)


S=V%*%t(V)
```


Joint sparse for each row of $V$:

```{r, fig.width=4, fig.height=4, echo=FALSE}
print(V)
image(as(V, "dgCMatrix"),main="V")
```

Let the covariance matrix be $$S=VV^\top+I$$
```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(V%*%t(V)+Ip, "dgCMatrix"),main="S=VV'+I")
```


Meanwhile the inverse of $S$ is


```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(solve(S+Ip), "dgCMatrix"),main="S inverse")
```

# 3. Estimation 

```{r,include=FALSE}
tridiag <- function(upper, lower, main){
    out <- matrix(0,length(main),length(main))
    diag(out) <- main
    indx <- seq.int(length(upper))
    out[cbind(indx+1,indx)] <- lower
    out[cbind(indx,indx+1)] <- upper
    return(out)
}
```


```{r}
p=15
r=0.5
V=matrix(0,p,p)
for(i in 1:p){
  for(j in 1:p){
    if((i==j) | (i-j)==2){
      V[i,j]=r
    }else if(i-j==1){
      V[i,j]=1
    }
  }
}


V[10:12,1]=c(r,1,r)

image(as(V, "dgCMatrix"),main="V")
S=V%*%t(V)+diag(rep(1,p))
image(as(S, "dgCMatrix"),main="S")
```


```{r,include=FALSE}
p=31
Ip=diag(rep(1,p))
V=matrix(0,p,29)
r=0.5
s=c(r,1,r)

for(i in 1:29){
  V[(i):(i+2),i]=s
}

p1=26
V[(p1-1):(p1+1),3]=s

S=V%*%t(V)+Ip

n=100
data = mvrnorm(n, mu = rep(0,p), Sigma = S)

```


```{r,include=FALSE}
M=tridiag(rep(1,p-1),rep(1,p-1),rep(1,p))
Shat=cov(data)
Stap=Shat*(abs(Shat)>0.6)
Stb=Shat*((abs(Shat)>0.6) | M)
```

```{r, include=F}
for(i in 2:(p-3)){
  for(j in (i+2):(p-1)){
    M0=Shat[(i-1):(i+1),(j-1):(j+1)]
    M[i,j]=M[j,i]=base::norm(M0,"F")>1.2
  }
}
Str=Shat*M
```



```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(V, "dgCMatrix"),main="V")
```


```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(S, "dgCMatrix"),main="S=VV'+I")
```

```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(Shat, "dgCMatrix"),main="Sample Covariance")
```

```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(Stap, "dgCMatrix"),main="Threshoulding")
```

```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(Stb, "dgCMatrix"),main="Threshoulding and banding")
```


```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(Str, "dgCMatrix"),main="Block Threshoulding and banding")
```



```{r, fig.width=4, fig.height=4, echo=FALSE}
image(as(solve(S), "dgCMatrix"),main="S inverse")
```



```{r}
base::norm(S-Shat,"F")
base::norm(S-Stap,"F")
base::norm(S-Stb,"F")
base::norm(S-Str,"F")
```

