---
title: "Covariance Model for Teleconnection"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source('../utils_cov.R')
```

# 1. NO teleconnection case 

```{r}
# True model (NO teleconnection)
p=100; r=0.5
V=genV(p=p,r=0.5,tele=F) # only neighbour effect
S=V%*%t(V)+diag(rep(1,p)) # true cov, generated from factor model S=VV'+I
printM(V)
printM(S)
```


## 1.0 sampling dis. of cov of two (univariate) ind. normal 

Given X, and Y are two independent normal (sample data with sample size 10^N2), what is the distribution of cov(X,Y)? (of course, the true value cov(X,Y)=0). This part simply repeat it N times, and find the sampling distribution of cov(X,Y). This is a preparation of the zero(true value) elements in the sample covariance matrix. 
 
```{r}
N=1000 # repeat size
N2=5 # sample size: 10 to the power
res=rep(0,N) # each iteration, save the covariance. all repeatants forms a sample
res2=rep(0,N2) # each sample size, save sd

for(j in 1:N2){
  n=10^j # sample size 
  for(i in 1:N){
    x=rnorm(n)
    y=rnorm(n)
    res[i]=cov(x,y)
  }
  res2[j]=sd(res) # standard error of the "sample covariance"
}

# sample size vs sd (of sampling dis.) and 1/sqrt(n)
plot(1:N2,log(res2),main="sd of sampling dis. of cov(true value = 0)", 
     xlab="log(sample size)",ylab="log(standard error)")
lines(1:N2,log(1/sqrt(10^(1:N2))),type="l",col="red")
legend("topright",c("log(sd)","log(1/sqrt(n))"),lty=c(1,1),pch = c(1,NA),col=c("black","red"))

# sampling distribution vs its normal version(same mean & sd)
plot(density(res),main="sampling dis. of cov ind. two normal")
x=seq(min(res),max(res),length=100)
y=dnorm(x,mean=mean(res), sd=sd(res))
lines(x,y, type="l", lwd=1,col="red")
legend("topright",c("sampling dis.","its normal appr."),lty=c(1,1),col=c("black","red"))
```


## 1.1 dis of zero/non-0s in sample cov 

```{r}
# (for one fixed sample cov) zero elements(off-diagonals) are the sample data
n=1000
data = mvrnorm(n, mu = rep(0,p), Sigma = S)
Shat=cov(data)

res=NULL
for(i in 1:p){
  for(j in 1:p){
    if((i-j)>2){
      res=c(res,Shat[i,j])
    }
  }
}

# sampling distribution vs its normal version(same mean & sd)
plot(density(res),col="blue",main="dis of zeros(true value) in sample cov")
x=seq(min(res),max(res),length=100)
y=dnorm(x,mean=mean(res), sd=sd(res))
lines(x,y, type="l", lwd=1,col="red")
legend("topright",c("dis. of zeros","its normal appr."),lty=c(1,1),col=c("black","red"))
```


## 1.2 dis of max zero/non-0 in many sample cov 

```{r}

N=1000
res2=rep(0,N)
for(k in 1:N){
  data = mvrnorm(n, mu = rep(0,p), Sigma = S)
  Shat=cov(data)
  # for(i in 1:p){
  #   Shat[i,i]=0
  # }
  # for(i in 1:(p-1)){
  #   Shat[i,i+1]=Shat[i+1,i]=0
  # }
  # for(i in 1:(p-2)){
  #   Shat[i,i+2]=Shat[i+2,i]=0
  # }
  res2[k]=max(abs(Shat*(S==0)))
}

plot(density(res2),col="blue",main="dis of max of zeros in many sample cov")
x=seq(min(res2),max(res2),length=100)
y=dnorm(x,mean=mean(res2), sd=sd(res2))
lines(x,y, type="l", lwd=1,col="red")
abline(v=mean(res2),lty=2,col="orange")
legend("topright",c("dis. of max of zeros","its normal appr."),lty=c(1,1),col=c("blue","red"))

summary(res2)
quantile(res2,0.99)
```


# 2. WITH teleconnection case 

```{r}
# True model (WITH teleconnection)
p=100; n=1000; r=0.9; s=0.5; tele=1; sigma=1
Ip=diag(rep(1,p))
V=genV(p=p,r=r,s=s,tele=tele)
S=V%*%t(V)+diag(rep(1,p))
M=(S==0)
printM(S)
```


## 2.1 one sample cov, distr of zero elements 

```{r}
n=1000
data = mvrnorm(n, mu = rep(0,p), Sigma = S)
Shat=cov(data)
# take zero elements in true cov
res=NULL

for(i in 1:p){
  for(j in 1:p){
    if(M[i,j] & (i-j>2)){
      res=c(res,Shat[i,j])
    }
  }
}

# result: dis. of zeros like normal
plotDen(res) 
```


## 2.2 multi sample cov, distr of max elements 

compare the distribution  of:

- max of zeros in true cov (Well, I exclude the margin 2<|i-j|<5, this part maybe large in the 3 by 3 block version)
- max of teleconnection signal (max off-diagonal non-zero center)

type: 0: sample cov; 1: spectual norm; 2: Frobenius norm; 3: mean; 4: abs(mean); 5: prod; 6: 2nd largest; 8: abs(prod(m0)); 10: sd; 11: skewness; 12: kurtosis;

r: rho; s: scale of the teleconnection s*(rho,1,rho); 

```{r}
# rho=1
p=100; n=1000; r=0.9; s=0.4; tele=1; sigma=1

par(mfrow=c(2,3))
for(i in c(0,1,2,3,4,10)){
  maxnoise(p,type=i, n=1000, N=100,r=r,s=s,tele=tele, sigma=sigma)
}
```

```{r}
# rho=-1
p=100; n=1000; r=0.9; s=0.4; tele=-1; sigma=1

par(mfrow=c(2,3))
for(i in c(0,1,2,3,4,10)){
  maxnoise(p,type=i, n=1000, N=100,r=r,s=s,tele=tele, sigma=sigma)
}
```


## 2.3 covariance estimation 


```{r}
# rho=1
p=100; n=1000; r=0.9; s=0.4; tele=1; sigma=1
Ip=diag(rep(1,p))
V=genV(p=p,r=r,s=s,tele=tele)
S=V%*%t(V)+diag(rep(1,p))

# Step I: Generate data 
# method 1: generate data directly with sigma
data = mvrnorm(n, mu = rep(0,p), Sigma = S)
# method 2: generate according to factor model
V=genV(p=p,r=r,s=s,tele=tele)
f=mvrnorm(n, mu = rep(0,p), Sigma = Ip)
E=mvrnorm(n, mu = rep(0,p), Sigma = Ip)
data = V%*%t(f)+sigma*t(E) #mvrnorm(n, mu = rep(0,p), Sigma = S)
data=t(data)

# Step II: Calculate Different Covariance Estimator
# NOTE:  each one has a threshould to tune based on different model

# TYPE 1: sample covariance
Shat=cov(data) 
# TYPE 2: (universal) threshoulding sample covariance
ST=Shat*(abs(Shat)>0.38) 
# note: large sample size n=1e5, thre=0.1, frob_thre=0.2, could find all, better than threshoulding (not not very interesting)
# TYPE 3: use (3 by 3 block) variance threshoulding
Snorm=MnormM(Shat,h=2,type=4,RM=F)
B1=(Snorm>0.38)
B2=matrix(0,p,p)
for(i in 1:p){
  for(j in 1:p){
    if(abs(i-j)<=10){
      B1[i,j]=0
    }
    if(abs(i-j)<=2){
      B2[i,j]=1
    }
  }
}
# SB=Shat*(B1 | B2) # B1: teleconnection; B2: banding, cut nearest-neighbours (too good)
SB=Shat*(Snorm>0.26)
# TYPE 4: Tony Cai's Adaptive threshould
SAT=MnormMvar(data) # Calculate variance of covariance elements
SA=regAT(Shat,sqrt(SAT),n,const=1.47)

printM(Shat)
printM(ST)
printM(SB)
printM(SA)
printM(S)
base::norm(S-Shat,"F")
base::norm(S-ST,"F")
base::norm(S-SB,"F")
base::norm(S-SA,"F")

round(S[4:8,89:93],2)
round(Shat[4:8,89:93],2)
round(Snorm[4:8,89:93],2)
```


```{r}
# rho=-1
p=100; n=1000; r=0.9; s=0.4; tele=-1; sigma=1
Ip=diag(rep(1,p))
V=genV(p=p,r=r,s=s,tele=tele)
S=V%*%t(V)+diag(rep(1,p))

# Step I: Generate data 
# method 1: generate data directly with sigma
data = mvrnorm(n, mu = rep(0,p), Sigma = S)
# method 2: generate according to factor model
V=genV(p=p,r=r,s=s,tele=tele)
f=mvrnorm(n, mu = rep(0,p), Sigma = Ip)
E=mvrnorm(n, mu = rep(0,p), Sigma = Ip)
data = V%*%t(f)+sigma*t(E) #mvrnorm(n, mu = rep(0,p), Sigma = S)
data=t(data)

# Step II: Calculate Different Covariance Estimator
# NOTE:  each one has a threshould to tune based on different model

# TYPE 1: sample covariance
Shat=cov(data) 
# TYPE 2: (universal) threshoulding sample covariance
ST=Shat*(abs(Shat)>0.38) 
# note: large sample size n=1e5, thre=0.1, frob_thre=0.2, could find all, better than threshoulding (not not very interesting)
# TYPE 3: use (3 by 3 block) variance threshoulding
Snorm=MnormM(Shat,h=2,type=4,RM=T)
B1=(Snorm>0.35)
B2=matrix(0,p,p)
for(i in 1:p){
  for(j in 1:p){
    if(abs(i-j)<=10){
      B1[i,j]=0
    }
    if(abs(i-j)<=2){
      B2[i,j]=1
    }
  }
}
# SB=Shat*(B1 | B2) # B1: teleconnection; B2: banding, cut nearest-neighbours (too good)
SB=Shat*(Snorm>0.26)
# TYPE 4: Tony Cai's Adaptive threshould
SAT=MnormMvar(data) # Calculate variance of covariance elements
SA=regAT(Shat,sqrt(SAT),n,const=1.47)

printM(Shat)
printM(ST)
printM(SB)
printM(SA)
printM(S)
base::norm(S-Shat,"F")
base::norm(S-ST,"F")
base::norm(S-SB,"F")
base::norm(S-SA,"F")

round(S[4:8,89:93],2)
round(Shat[4:8,89:93],2)
round(Snorm[4:8,89:93],2)
```




